---
title: "Bar Graph 1: Landscape of Arts Education Nonprofits"
author: "Steven Todd McIntyre II"
date: "2026-02-23"
runtime: shiny
format: 
  dashboard:
    embed-resources: true
    theme: minty
    output:
      orientation: columns
execute:
  warning: false
  message: false
  echo: false
---

```{r}
library(tidyverse)
library(tidytext)
library(here)
library(knitr)
library(shiny)
library(rio)
library(plotly)

```

# Overview

## Column
**How useful can NTEE Codes be in Studying Nonprofits?**

This dashboard will explore United States nonprofit organizations that are coded as "A25" in the National Taxonomy of Exempt Organizations (NTEE), a coding scheme that describes kinds of nonprofit organizations. "A25" is defined by the Urban Institute as "Organizations that provide informal arts educational programming and/or instruction but do not grant diplomas or degrees; or which offer services regarding the arts to educational institutions or to public entities involved in education." Using data from the [Urban Institute's Business Master File (BMF)](https://urbaninstitute.github.io/nccs/datasets/bmf/), I explore the kinds of organizations that use A25 

# Visualization 1: A25 Word Count

## Column

```{r}
a25 <- import(here("data/BMF_A25.csv"))

kind_of_org <- c("foundation")
art_words <- c("arts", "art")
target_audience <- c("artists", "american", "community", "youth")

a25_top_50 <- a25 |>
  unnest_tokens(output = words, input = ORG_NAME_CURRENT) |>
  select(words) |>
  filter(!words %in% stop_words$word) |>
  count(words) |>
  mutate(words = fct_lump_n(words, n=50, w=n)) |>
  filter(words != "Other") |>
  group_by(words) |>
  summarize(n = sum(n)) |>
  mutate(words=fct_reorder(words, n))

a24_top_50_plot <- a25_top_50 |>
  ggplot(aes(n, words)) +
  geom_col() +
  geom_col(aes(y=filter(a25_top_50, words=="arts")$words), fill = "red") +
#  geom_col(aes(y=filter(a25_top_50, words%in%target_audience)$words), fill="blue") +
  labs(
    title = "Top 50 Words in A25 \"Arts Education\" Organization Names",
    x = "Instances in A25 Organization Names",
    y = "",
    caption = "Data from the Urban Institute's Business Master File."
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 4)
  )

ggplotly(a24_top_50_plot)
```

# Summary of Updates

## Column

Removing the relatively poor visualization that was mostly intended to test embed-resources, I've now added a visualization that will be, in some form, in my final project.

Welcome!

Today, I took the following steps:

-   I created this blog!
-   I cleaned and uploaded data from the [Business Master File (BMF)](https://urbaninstitute.github.io/nccs/datasets/bmf/), including only data with an A25 (Arts Education) NTEE (National Taxonomy of Exempt Entities) code.
-   I made a pretty bad visualization in order to test embed-resources
