---
title: "Labels and Colors"
author: "Steven Todd McIntyre II"
date: "2026-03-10"
runtime: shiny
format: 
  dashboard:
    embed-resources: true
    theme: minty
    output:
      orientation: columns
execute:
  warning: false
  message: false
  echo: false
---

```{r}
library(tidyverse)
library(tidytext)
library(here)
library(knitr)
library(shiny)
library(rio)
library(plotly)
library(usmap)
library(scales)
library(ggrepel)
library(ggbeeswarm)

```

# Overview

## Column

**How useful can NTEE Codes be in Studying Arts Education Nonprofits?**

This dashboard will explore United States nonprofit organizations that are coded as "A25: Arts Education" in the National Taxonomy of Exempt Organizations (NTEE), a coding scheme that describes kinds of nonprofit organizations. "A25" is defined by the Urban Institute as "Organizations that provide informal arts educational programming and/or instruction but do not grant diplomas or degrees; or which offer services regarding the arts to educational institutions or to public entities involved in education." Using data from the [Urban Institute's Business Master File (BMF)](https://urbaninstitute.github.io/nccs/datasets/bmf/), I explore the kinds of organizations that define themselves as A25 and consider the NTEE classification's usefulness in exploring nonprofit arts education activities.

# Visualization 1: Arts Education Nonprofit Title Word Count

## Column

```{r}
a25 <- import(here("data/BMF_A25.csv"))

# set categories of top 50 words
kind_of_org <- c("foundation", "institute", "school", "center", "academy", "association", "guild", "society", "project", "studio", "incorporated", "club", "corporation", "alliance", "productions", "program", "council", "collective", "nfp", "league", "friends", "corp", "boosters", "gallery")
art_words <- c("arts", "art", "music", "creative", "fine", "design", "theatre", "performing", "workshop", "graphic", "dance", "traditional", "cultural", "hooking", "theater", "quilters", "woodturners")
education_words <- c("education", "learning", "educational")
locations <- c("american", "community", "international", "county", "valley", "city")
target_audience <- c("artists", "youth", "kids", "artist", "childrens")

# remove orgs that no longer exist
a25 <- a25 |>
  filter(ORG_YEAR_LAST >= 2024)

# create a table of the top 50 words in org names
a25_top_50 <- a25 |>
  unnest_tokens(output = word, input = ORG_NAME_CURRENT) |>
  select(word) |>
  filter(!word %in% stop_words$word) |>
  count(word) |>
  mutate(word = fct_lump_n(word, n=50, w=n)) |>
  filter(word != "Other") |>
  group_by(word) |>
  summarize(n = sum(n)) |>
  mutate(word=fct_reorder(word, n))

# add categories to the top 50 words
a25_top_50 <- a25_top_50 |>
  mutate(category = case_when(
    word%in%kind_of_org ~ "organization",
    word%in%art_words ~ "art",
    word%in%education_words ~ "education",
    word%in%locations ~ "location",
    word%in%target_audience ~ "people"
  ))

# create plot
a25_top_50_plot <- a25_top_50 |>
  ggplot(aes(n, word, fill = category)) +
  geom_col(width = .9, position = position_dodge(.7)) +
  labs(
    title = "Top 50 Words in Arts Education Nonprofit Names",
    x = "Instances in Organization Names",
    y = "",
    fill = "Word Category",
    caption = "Data from the Urban Institute's Business Master File. Categories coded by author."
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8) 
  ) +
  scale_fill_discrete(palette = "Set2")

ggplotly(a25_top_50_plot)
```

# Visualization 2: US Map

## Column

```{r}
# create table with count of A25 in each state
a25_state_count <- a25 |>
  count(F990_ORG_ADDR_STATE) |>
  mutate(state = F990_ORG_ADDR_STATE) |>
  select(state, n) |>
  filter(state != "")

# create map plot using US Map
a25_dist_map <- plot_usmap(data = a25_state_count, values = "n") +
  labs(title = "Where are the Arts Education Nonprofits?",
       subtitle = "Count of arts education nonprofits in each US state.",
       fill = "Number of Arts\nEducation Nonprofits",
       caption = "Data from the Urban Institute's Business Master File.") +
  theme_minimal() +
  scale_fill_continuous(palette = "PuBu")

ggplotly(a25_dist_map)

```

# Visualization 3: Arts Education Nonprofits vs. State Population

## Column {.tabset}

```{r}
#| title: "Static, With Labels"
# join usmaps' state population table w/ a25 count data
a25_statepop <- inner_join(a25_state_count, statepop, by = join_by(state == abbr))

# create scatterplot with fit line, highlight higher and lower states
a25_statepop_scatter <- a25_statepop |>
  ggplot(aes(x=pop_2022, y=n, label=state)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey90") +
  geom_point(color = "grey60") +
  geom_point(data = a25_statepop |> filter(state%in%c("CA", "NY", "OR", "MA", "VT", "DC")), color = "green") +
  geom_point(data = a25_statepop |> filter(state%in%c("PR", "MS", "SC", "NE", "OK")), color = "red") +
  geom_text_repel(color = "black") +
  theme_minimal() +
  labs(
    title = "Variation of Arts Education Nonprofits vs. State Population",
    subtitle = "Count of arts education nonprofits in each US state. Note states above and below the line. Any trends?",
    x = "2022 State Population (log10)",
    y = "Number of Arts Education Nonprofits (log10)",
    caption = "Nonprofit data from the Urban Institute's Business Master File. State population data provided through \`usmap\` package.\nEach are plotted logarithmically for better spacing between lower-populated states."
    
  ) +
  scale_color_manual(values = c()) +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10()

a25_statepop_scatter
```

```{r}
#| title: "Interactive, No Labels"
#make ggplotly version
ggplotly(a25_statepop_scatter)
```


# Visualization 4: 2024 Arts Education Nonprofit Revenues

## Column {.tabset}
```{r}
#| title: "Log10 Distribution"
# Make Distribution Beeswarm Plot
a25_rev_distr_log <- a25 |>
  ggplot(aes(y="", x = F990_TOTAL_REVENUE_RECENT, label = ORG_NAME_CURRENT)) +
  geom_quasirandom(color = "darkgrey") +
  scale_x_log10(labels = scales::label_currency(prefix = "$", big.mark = ",")) +
  theme_minimal() +
  labs(
    title = "2024 Revenue Distribution for Arts Education Nonprofits",
    y = "",
    x = "Total Revenue Reported in 2024 F990 Form (log10)",
    caption = "Nonprofit data from the Urban Institute's Business Master File."
  )

ggplotly(a25_rev_distr_log)
```

```{r}
#| title: "Linear Distribution"
a25_rev_distr_lin <- a25 |>
  ggplot(aes(y="", x = F990_TOTAL_REVENUE_RECENT, label = ORG_NAME_CURRENT)) +
  geom_quasirandom(color = "darkgrey") +
  scale_x_continuous(labels = scales::label_currency(prefix = "$", big.mark = ",")) +
  theme_minimal() +
  labs(
    title = "2024 Revenue Distribution for Arts Education Nonprofits",
    y = "",
    x = "Total Revenue Reported in 2024 F990 Form",
    caption = "Nonprofit data from the Urban Institute's Business Master File."
  )

ggplotly(a25_rev_distr_lin)
```

# Summary of Updates

## Column

This pass, in preparation for my presentation, was for labels, colors, and readability pieces. Here's to trying to tell better data stories!

In this post, I did the following:

-   V1: Changed colors, edited labels.
-   V2: Added/edited labels.
-   V3: Added labels. Added a label for each state using ggrepel. Placed each axis on a logarithmic scale for a better viewing experience.
-   V4: Changed from a boxplot to a beeswarm plot in ggplotly, to examine outliers while showing overall distribution. Retained log10 x-axis. Added labels and tabset with two versions present.
